\bf What is the Best Way to Display Two Time Series to Read Association?

### 2017-01-02 Confirmation/confirmation_outline.tex
Xmas break updates

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

@@

\usepackage{alltt}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\numberwithin{equation}{section}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\tableofcontents

\section{Bayesian Inference}

\subsection{Bayes' Rule}

In Bayesian statistics, we are interested in making inferences about an unknown parameter $\theta$ using what is known as the posterior distribution, $p(\theta | y)$, where $y$ is our observed data. Bayes' rule provides a form for this distribution,

\begin{equation}
\label{BR1}
p(\theta | y) = \frac{p(y | \theta) p(\theta)}{p(y)},
\end{equation}

which shows that the posterior can be found by taking the product of the likelihood, $p(y | \theta)$ and the prior distribution, $p(\theta)$, and normalising with a constant $p(y)$. In practice $p(y)$ is unknown and can be found from

\begin{equation}
\label{BR2}
p(y) = \int_\theta p(y | \theta) p(\theta) d \theta.
\end{equation}

Finding the value of $p(y)$ is often the most difficult part of working with Bayesian statistics. In some simple models, this integration can be solved analytically, but for most models of interest it is intractable and we can only work with the un-normalised version

\begin{equation}
\label{BR3}
p(\theta | y)^* = p(y | \theta) p(\theta).
\end{equation}

This work will concern itself with two of the methods used in the case where (\ref{BR2}) is intractable, exact methods such as MCMC, which will be discussed later in this section as well as an approximation method known as Variational Bayes which appears in section two.

\subsection{Exact Bayesian Computation}

\textbf{1. Numerical Integration}\newline

It might be tempting to numerically approximate (\ref{BR2}) by evaluting the integrand over a grid and applying methods such as the Trapezoid Rule or Adaptive Quadrature. These may be suitable for integration in a low dimension, but if we have a grid with $G$ points in one dimension, then a two dimensional integration requires $G^2$ points, and a three dimensional integration requires $G^3$ points. In general, the computational time increases exponentially with the dimension of $\theta$ in numerical integration. In almost all cases of interest, $\theta$ is a high-dimensional vector which makes numerical integration prohibitively expensive. This is known as the curse of dimensionality, where the difficulty of the problem increases quickly with the dimension of $\theta$ and requires us to use more sophisticated methods.\newline

\textbf{2. Importance Sampling} \newline

Candidate Density
Accept Reject
Good candidate is hard but useful
VB as a candidate \newline

\textbf{3. Gibbs Sampling}\newline

Breaks down into simple problems
Becomes linear instead of exponential
Exact sampling \newline

\textbf{4. Metropolis Hastings within Gibbs}\newline

Gibbs only works exactly on some problems
Approximation within Gibbs
Also can benefit from VB

\section{Variational Bayes}

\subsection{Estimation as an optimisation problem}

In the event that our dataset is too large, or our model is too complex, for MCMC to converge within a reasonable time frame we must resort to approximation. One important question that might be asked then is \textit{`How do I measure how good my approximation is?'}. If two distributions come from the same family, one might consider the difference in the parameters, but it is trivial to come up with distributions with slight parameter differences and vastly different regions of probability mass. Consider the difference between a $\mathcal{N}(0, 0.1)$ and a $\mathcal{N}(0.5, 0.1)$ distribution, the change in mean is only $0.5$ but there is little overlap in the densities. Further, large values for the variance allow large changes in mean with very little change in the resulting densities. If the distributions come from different families it is almost impossible to use parameters to measure the degree of closeness between two distributions. Instead we must move from the parameter space to the distribution space, where one measure of the difference between distributions $p(\theta)$ and $q(\theta)$ is the Kullback-Leibler (KL) divergence \citep{Kullback1951}, defined by

\begin{equation}
\label{KL-def}
KL[q(\theta)||p(\theta)] = \int q(\theta) \ln \left( \frac{q(\theta)}{p(\theta)}\right) d\theta.
\end{equation}

KL divergence is a non-negative, assymetric measure of the difference between $p(\theta)$ and $q(\theta)$ that will equal zero if and only if $p(\theta) = q(\theta)$ almost everywhere. It has origins in information theory, and can be interpreted as: \textit{`Given that I know $p(\theta)$ for some $\theta \in \Theta$, how much extra information is required, on average, to know the value of $q(\theta)$?'} There are examples of the literature of approximations with other measures of divergence, such as \citet{Minka2001} introducting Expectation Propogation, which minimises the reverse measure $KL[p(\theta)||q(\theta)]$, which was extended to Power-EP in \citet{Minka2004}, which aims to minimisea the more general $\alpha-\mbox{divergence}$ \citep{Amari1985}.
\vspace{3mm}

\begin{figure}[hb]
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/klnorm-1}

\end{knitrout}
\caption{Red: A Normal(0, 0.1) distribution. Green: A Normal(0.5, 0.1) distribution.}
\end{figure}

\newpage

Variational Bayes aims to find the approximating distribution, $q(\theta | \lambda)$, that minimises $KL[q(\theta | \lambda)||p(\theta |y)]$. This approximating distribution is parameterised by a $\lambda$ vector, each element of which may be a function of $y$, the hyperparameters of the prior distribution, and even of other elements in $\lambda$. The optimal approximation, in the sense of having the minimum KL divergence with the posterior, can be derived by beginning with the unknown constant, $\ln(p(y))$.
\begin{align}
\ln(p(y)) &= \int_{\theta} q(\theta|\lambda) \ln(p(y)) d\theta \nonumber \\
&= \int_{\theta} q(\theta|\lambda) \ln  \left( \frac{p(y, \theta)}{p(\theta | y)} \right) d\theta \nonumber \\
&= \int_{\theta} q(\theta|\lambda) \ln  \left( \frac{p(y, \theta)}{p(\theta | y)} \frac{q(\theta|\lambda)}{q(\theta|\lambda)}\right) d\theta \nonumber \\
&= \int_{\theta} q(\theta|\lambda) \ln  \left( \frac{q(\theta|\lambda)}{p(\theta | y)}\right) + \ln \left( \frac{p(y, \theta)}{q(\theta|\lambda)} \right) d\theta \nonumber \\
&= \int_{\theta} q(\theta|\lambda) \ln  \left( \frac{q(\theta|\lambda)}{p(\theta | y)}\right) d\theta + \int_{\theta} q(\theta|\lambda) \ln \left( \frac{p(y, \theta)}{q(\theta|\lambda)} \right) d\theta \nonumber \\
&= KL [q(\theta||\lambda) || p(\theta | y)] + \mathcal{L}(q(\theta | \lambda), y) \nonumber
\end{align}

where

\begin{align}
\mathcal{L}(q(\theta|\lambda), y) &= \int_{\theta} q(\theta|\lambda) \ln \left( \frac{p(y, \theta|\lambda)}{q(\theta|\lambda)} \right) d\theta \nonumber \\
&=  \int_{\theta} q(\theta|\lambda) \ln (p(y, \theta|\lambda)) d\theta -  \int_{\theta} q(\theta|\lambda) \ln (q(\theta|\lambda)) d\theta. \label{ELBO}
\end{align}

Hence we have that
\begin{equation}
\label{VBKL}
KL [q(\theta|\lambda) || p(\theta | y)] = \ln(p(y)) - \mathcal{L}(q(\theta | \lambda), y).
\end{equation}

From (\ref{VBKL}) it is clear that by maximising $\mathcal{L}(q(\theta | \lambda), y)$ with respect to $q(\theta|\lambda)$ the KL divergence between the unknown true posterior $p(\theta|y)$ and the approximating distribution $q(\theta|\lambda)$ will be minimised.
Hence the Variational Bayes solution will be to maximise the Evidence Lower Bound (ELBO), $\mathcal{L}(q(\theta|\lambda), y)$, in (\ref{ELBO}).

\subsection{Mean Field Variational Bayes}

@@

If the likelihood and prior for $\theta_i$ form an exponential family conjugate pair, then this maximisation proceeds by keeping the distributional family constant and simply updating the parameters $\lambda_i$. Otherwise, we can find the optimal distributional family for any $q_i$ by taking the log-joint density and then ignoring any term that doesn't depend on $\theta_i$. Next, take the expectation of any remaining $\theta_{j \neq i}$ and then match to the kernel of a known distribution. If it does not match a known distribution, we may need to make a further approximation $\tilde{q_i(\theta_i|\lambda_i)}$ which has a recognizable distribution. One method as used in \citet{Friston2006} uses a Laplace approximation for the otherwise unrecognizable $q_i(\theta_i | \lambda_i)$ with reasonable results.

These steps are almost the same as those used to find a conditional distribution $p(\theta_i | \theta_{j \neq i}, y)$, with dependence on other parameters replaced by their expectations. Hence, the optimal approximating family for $\theta_i$, $q(\theta_i | \lambda_i)$, will come from the same distributional family as the conditional distribution $p(\theta_i | \theta_{j \neq i}, y)$, if it exists in a recognisable form such as the exponential family. Deriving the updating algorithm for MFVB takes more analytical effort than finding the conditionals for a Gibbs MCMC scheme, as we must then find the values of each $\lambda_i$ implied by (\ref{mf3}), which are often a function of $\lambda_{j \neq i}$ through the expectations.
\vspace{5mm}

Given these optimal distributional families, an algorithm that cycles through all of the $\lambda_i$ and sets each to its maximising value is known as a coordiante ascent algorithm. The algorithm must continuously iterate between each parameter until $\mathcal{L}(q(\theta | \lambda), y)$ converges with some pre-defined threshold noting that the $\lambda$ may depend on each other. If $\lambda$ has dimension $k$, the algorithm follows below.

@@

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Log Joint Density}
 \KwResult{Mean Field Approximation}
 Initialise $\lambda$ randomly\;
 \While{Not converged}{
  \For{$i =  1$ \KwTo $k$}{
      Hold $\lambda_{j \neq i}$ fixed\;
      Use (\ref{mf3}) to match $q(\theta_i|\lambda_i)$ to a known distribution\;
      Update $\lambda_i$ using the most recent values of $\lambda_{j \neq i}$\;
     }
 }
 \caption{Coordinate Ascent for MFVB}
  \label{alg:algorithm1}
\end{algorithm}

\subsection{Stochastic Gradient Ascent}

In many cases, the posterior is too complex to be captured by a factorising approximation and we desire our distribution $q(\theta | \lambda)$ to capture dependence between parameters. In this case, the easy maximisation in MFVB is unavailable, and we must resort to what is called Stochastic Variational Bayes (SVB) using a gradient ascent algorithm developed by \citet{Paisley2012} and \citet{Hoffman2013}.

@@

\subsection{SVB Extensions}

\textbf{1. AdaGrad} \newline

If the gradient is small, we want the learning rate, $\rho_t$, to be large, and if the gradient is large, we want the learning rate to be small. AdaGrad \citep{Duchi2011} is a method of setting the learning rate to dynamically adapt as the values of the calculated gradients changes. Further, AdaGrad allows the step size of each element of $\lambda$ to move independently of the other elements of $\lambda$.

If $\nabla_{\lambda}^{(m)}$ is the vector of gradients at iteration $m$, define

\begin{equation}
G^{(m)} = \sum_{i = 1}^{m} \nabla_{\lambda}^{i} (\nabla_{\lambda}^{i})^T,
\end{equation}

the sum of the outer products of the gradients up to iteration $m$. Then set

\begin{equation}
\label{adagrad}
\rho^{(m)} = \eta \mbox{diag}(G^{(m)})^{-1/2}
\end{equation}

with some preset tuning parameter $\eta$. If we have $k$ elements in $\lambda$, $G^{(m)}$ is a $k \times k$ matrix so we have one element in $\rho^{(m)}$ for each element of $\lambda$. As we only need the diagonal of $G^{(m)}$, this may be calculated in linear time. \newline

\textbf{2. Reparameterisation} \newline

The reparameterisation trick was popularised following \citet{Kingma2013} and allows the gradient of the ELBO to be simplified.

Consider a parameter free auxillary distribution $q(\epsilon)$ and differentiable transformation $f(\cdot,\cdot)$ such that $\theta = f(\epsilon, \lambda)$. Examples include a location-scale transformation from a standard normal or an inverse-CDF transform from a uniform$(0, 1)$ variable. Note that

$$q_\theta(\theta | \lambda) = q_\epsilon(\epsilon) \left| \frac{d\epsilon}{d\theta} \right| $$

where the parameters that govern the transform $f$ are the same $\lambda$ parameters as in $q(\theta | \lambda)$. Now (\ref{SGA2}) becomes

\begin{align}
\nabla_{\lambda} \mathcal{L}(q(\theta | \lambda), y) &=  \int_{\epsilon} \nabla_{\lambda}[  q(\epsilon) \left( \ln (p(y, f(\epsilon, \lambda))) - \ln(q(f(\epsilon, \lambda) | \lambda) \right)] d\epsilon \nonumber \\
&=  \int_{\epsilon}  q(\epsilon) \nabla_{\lambda} \left[ \ln (p(y, f(\epsilon, \lambda))) - \ln(q(f(\epsilon, \lambda)| \lambda) \right] d\epsilon \label{RP1} \\
&= \int_{\epsilon}  q(\epsilon) \nabla_{\lambda} \left[ \ln (p(y, f(\epsilon, \lambda)) \right] d\epsilon\nonumber
\end{align}

as the second part on the right hand side of (\ref{RP1}) is a score function with expectation equal to zero. This can be estimated using

\begin{equation}
\label{RP2}
 \mathcal{L}(q(\theta | \lambda), y) \approx  \frac{1}{S}\sum_{s=1}^{S} \nabla_{\lambda} \left[ \ln (p(y, f(\epsilon_s, \lambda))) \right].
\end{equation}

with $\epsilon_s \sim q(\epsilon)$.

\citet{Kingma2015} explore how this reparameterised version can have orders of magnitude lower variance than the estimator in (\ref{SGA5}).

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Log Joint Density, Approximation family q}
 \KwResult{Variational Approximation}
 Initialise $\lambda$\;
 \While{Not converged}{
  Simulate $\epsilon^s$ for $s = 1, \dots S$\;
  \For{$i =  1$ \KwTo $k$}{
      Calculate $\nabla_{\lambda_i} =  1/S\sum_{s=1}^{S} \nabla_{\lambda_i}[\log(p(f(\epsilon^s, \lambda^{(m-1)}), y))]$
     }
  Set $\lambda^{(m)} = \lambda^{(m-1)} + \rho^{(m)} \nabla_{\lambda}^{t}$\;
  Set $m = m + 1$\;
 }
 \caption{Gradient Ascent for re-parameterised SVB}
  \label{alg:algorithm3}
\end{algorithm}

$\mbox{ }$\newline \textbf{3. Control Variates} \newline

If reparameterisation is too difficult or otherwise unavailable, \citet{Paisly2012} introduce the use use of control variates to estimate the gradient with reduced variance.

Define

\begin{equation}
\label{CV1}
g(\theta, \lambda, y) =  \nabla_{\lambda} [\ln(q(\theta_s | \lambda))] (\ln (p(y, \theta_s)) - \ln(q(\theta_s | \lambda))),
\end{equation}

the function estimated in (\ref{SGA5}), then for some other function $h$, define

\begin{equation}
\label{CV2}
\hat{g}(\theta, \lambda, y) = g(\theta, \lambda, y) - a(h - \mathbb{E}(h)).
\end{equation}

Then we have that

\begin{equation}
\mathbb{E}(\hat{g}) = \mathbb{E}(g)
\end{equation}

\begin{equation}
\label{CV3}
\mbox{Var}(\hat{g}) = \mbox{Var}(g) + a^2 \mbox{Var}(h) - 2a\mbox{Cov}(g, h). 
\end{equation}

Solving the polynomial (\ref{CV3}) in $a$ shows that the variance of $\hat{g}$ is minimised by

\begin{equation}
\label{CV4}
\hat{a} = \mbox{Cov}(g, h)/\mbox{Var}(h).
\end{equation}

Substituting (\ref{CV4}) into (\ref{CV3}) yields

\begin{equation}
\label{CV5}
\mbox{Var}(\hat{g}) = \mbox{Var}(g) - \mbox{Cov}(g, h)^2/\mbox{Var}(h).
\end{equation}

We need to choose some function $h$ that has a large covariance with $g$. \citet{Ranganath2014} suggests that the an easy option is

\begin{equation}
\label{CV6}
h(\theta, \lambda) = \nabla_{\lambda} [\ln(q(\theta_s | \lambda))].
\end{equation}

As $h$ is a score function, it has an expectation of zero and our estimator becomes

\begin{equation}
\label{CV7}
\nabla_{\lambda} \mathcal{L}(q(\theta | \lambda), y) \approx \frac{1}{S}\sum_{s=1}^{S} \nabla_{\lambda} [\ln(q(\theta_s | \lambda))] (\ln (p(y, \theta_s)) - \ln(q(\theta_s | \lambda)) - \hat{a}),
\end{equation}

where $\hat{a}$ is estimated from the sample variance and covariance of a subset of the $S$ Monte-Carlo draws.

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Log Joint Density, Approximation family q}
 \KwResult{Variational Approximation}
 Initialise $\lambda$\;
 \While{Not converged}{
  Simulate $\theta_s$ for $s = 1, \dots S$ from $q(\theta|\lambda^{(m-1)})$\;
  \For{$i =  1$ \KwTo $k$}{
      Calculate $g_{\lambda_i} = 1/S\sum_{s=1}^{S} \nabla_{\lambda_i}[\log(q(\theta^s|\lambda^{(m-1)})](\log(p(\theta^s, y)) - \log(q(\theta^s |\lambda^{(m-1)})))$\;
      Calculate $h_{\lambda_i} = 1/S\sum_{s=1}^{S}\nabla_{\lambda_i}[\log(q(\theta^s |\lambda^{(m-1)}))]$\;
      Use a subset of $S$ to estimate $\hat{a} = \mbox{Cov}(h, g)/\mbox{Var}(g)$\;
      Calculate $\nabla_{\lambda_i} = g_{\lambda_i} - \hat{a} h_{\lambda_i}$\;
     }
  Set $\lambda^{(m)} = \lambda^{(m-1)} + \rho^{(m)}  \nabla_{\lambda}$\;
  Set $m = m + 1$\;
 }
 \caption{Gradient Ascent for SVB with control variates}
  \label{alg:algorithm4}
\end{algorithm}

\subsection{AR2 model example}

Consider $T$ observations from an AR(2) model conditioned on $y_1$ and $y_2$,

\begin{equation}
\label{AR2}
y_t = \phi_1 y_{t-1} + \phi_2 y_{t-2} + \epsilon_t \mbox{ for } t = 3,\dots,T,
\end{equation}

where $\epsilon$ are iid $\mathcal{N}(0, \sigma^2)$.

Using the independent priors:

\begin{align}
p(\phi_1) &\sim \mathcal{N}(\bar{\phi_1}, \tau_{1}) \nonumber \\
p(\phi_2) &\sim \mathcal{N}(\bar{\phi_2}, \tau_{2}) \nonumber \\
p(\sigma^2) &\sim \mathcal{IG} (\mbox{shape} = a, \mbox{ scale} = b), \nonumber
\end{align}

we have a model with zero stationary restrictions which can easily be sampled through a Gibbs MCMC scheme. This exact posterior can be used in comparision with the results of our variational approximations.

The Gibbs conditional distributions are derived in the appendix and are

\begin{align}
p(\phi_1 | \phi_2, \sigma^2, y_{1:T}) &\propto \mathcal{N}(\mu_1^G, \gamma_1^G) \nonumber \\
p(\phi_2 | \phi_1, \sigma^2, y_{1:T}) &\propto \mathcal{N}(\mu_2^G, \gamma_2^G) \nonumber \\
p(\sigma^2 | \phi_1, \phi_2, y_{1:T}) &\propto \mathcal{IG}(shape = \alpha^G,  scale = \beta^G), \nonumber
\end{align}

where

\begin{align}
\mu_1^G &= \frac{\tau_1 (\sum_{t=3}^{T} y_t y_{t-1} - \phi_2 \sum_{t=3}^{T} y_{t-1} y_{t-2}) + \sigma^2 \bar{\phi_1}}
          {\tau_1 \sum_{t=3}^{T} y_{t-1}^2 + \sigma^2} \nonumber \\
\gamma_1^G &= \frac{\sigma^2 \tau_1}{\tau_1 \sum_{t=3}^{T} y_{t-1}^2 + \sigma^2} \nonumber \\
\mu_2^G &= \frac{\tau_2 (\sum_{t=3}^{T} y_t y_{t-2} - \phi_1 \sum_{t=3}^{T} y_{t-1} y_{t-2}) + \sigma^2 \bar{\phi_2}}
          {\tau_2 \sum_{t=3}^{T} y_{t-2}^2 + \sigma^2} \nonumber \\
\gamma_2^G &= \frac{\sigma^2 \tau_2}{\tau_2 \sum_{t=3}^{T} y_{t-2}^2 + \sigma^2} \nonumber \\
\alpha^G &= a + T/2 - 1 \nonumber \\
\beta^G &= b + 1/2 \sum_{t=3}^{T} \left((y_t - \phi_1 y_{t-1} - \phi_2 y_{t-2})^2 \right).  \nonumber
\end{align}

100 data points are simulated from the model in (\ref{AR2}) with true values of $\phi_1 = 0.8, \phi_2 = 0.15$ and $\sigma^2 = 2$. These values were chosen to induce a strong dependence between the $\phi$ parameters as it is close to the non-stationary region of $\phi_1 + \phi_2 > 1$. The resulting posterior is sampled with 100,000 draws from Gibbs MCMC, discarding the first 20\% of all draws as a burn in. The resulting draws were strongly dependent so they were thinned by a factor of 20 to result in 4000 effectively independent draws. The total runtime for MCMC was 2.6 seconds. The dependence between $\phi_1$ and $\phi_2$ is apparent in figure (\ref{AR2MCMC}), so a SGA algorithm that captures this dependence should be a better approximation than a MFVB algorithm that assumes independence.

\begin{figure}
\label{AR2MCMC}
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/AR2MCMC-1}

\end{knitrout}
\caption{Marginal and bivariate density plots of the MCMC draws. 100,000 draws per parameter were sampled from (\ref{AR2}) via Gibbs MCMC then reduced to 4000 each. Note the strong dependence between the two $\phi$ parameters.}
\end{figure}

The derivation of the MFVB approximation is also in the appendix and each marginal has the same form as the corresponding true conditional distribution above,
\begin{align}
q(\phi_1 | \lambda) &\propto \mathcal{N}(\mu_1^{MF}, \gamma_1^{MF}) \nonumber \\
q(\phi_2 | \lambda) &\propto \mathcal{N}(\mu_2^{MF}, \gamma_2^{MF}) \nonumber \\
q(\sigma^2 | \lambda) &\propto \mathcal{IG}(shape = \alpha^{MF},  scale = \beta^{MF}), \nonumber
\end{align}

where the parameterising vector $\lambda = (\mu_1^{MF}, \gamma_1^{MF}, \mu_2^{MF}, \gamma_2^{MF}, \alpha^{MF}, \beta^{MF})'$.

The MVFB derivation yields the following updating equations for $\lambda$,

>>>>

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\numberwithin{equation}{section}
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

@@

\usepackage{alltt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\tableofcontents
\section{Introduction}

For many time-series applications a point forecast of the mean of the next observation in a series, $y_{T+1}$, can be easily obtained with frequentist methods, conditioning on the forecaster's observed series, $y_{1:T}$. Often the forecast is supplemented with a prediction interval backed by asymptotic theory, however there is a growing demand in the literature for density forecasts (find citations, probably Hyndman) which are typically much harder to obtain.

As a contrast to frequentism, the Bayesian methodology implicitly provides the entire probability density for variables of interest, in this case $p(y_{T+1} | y_{1:T})$. However the Bayesian approach requires solving integrals of the form

\begin{equation}
\label{forecastintegral}
\int_\theta p(y_{T+1} | \theta) p(\theta | y_{1:T}) d \theta,
\end{equation}

where $p(\theta | y_{1:T})$ is known as the posterior distribution, with

\begin{equation}
\label{posterior}
 p(\theta | y_{1:T}) = \frac{p(y_{1:T}|\theta)p(\theta)}{\int_\theta p(y_{1:T}|\theta)p(\theta) d\theta}
\end{equation}

and $p(\theta)$ is some pre-specified prior distribution.

These often cannot be solved analytically, while numeric integration is computationally infeasible when the dimension of $\theta$ is large. To address this problem there is a wide range of techniques used to approximate the solution to the integral in (\ref{forecastintegral}), such as Markov Chain Monte Carlo (MCMC) and Variational Bayes (VB). These approximations have an implicit trade-off: The better approximations are compuationaly intensive, and the forecaster must decide how much computation time and approximation error is acceptable. The focus of this research is in situtations where the time budget is too small for MCMC to be reliable, and uses VB as an alternative. This situtation is common in time-series forecasting, which must update the posterior distribution for each new data point observed via

\begin{equation}
\label{posteriorupdate}
p(\theta | y_{1:T+1}) = \frac{p(y_{T+1} | \theta) p(\theta | y_{1:T})}{\int_\theta p(y_{T+1} | \theta) p(\theta | y_{1:T}) d\theta}.
\end{equation}

\section{Bayesian Inference}
\subsection{Exact Bayesian Computation}
While it is technically an approximation method, MCMC algorithms result in what is often called an exact computation of the posterior. A Gibbs based MCMC iteratively samples from the conditional distributions

\begin{align}
&p(\theta_1 | \theta_2, \dots, \theta_p, y_{1:T}) \nonumber \\
&p(\theta_2 | \theta_1, \theta_3, \dots, \theta_p, y_{1:T}) \nonumber \\
&\vdots \nonumber \\
&p(\theta_p | \theta_1, \dots, \theta_{p-1}, y_{1:T}) \nonumber
\end{align}

where $p$ is the dimension of $\theta$. With enough iterations, the error in the approximation converges to zero and the algorithm can be ran for as much time as the forecaster desires to reduce error to a desired level. However, in the first iteration we must set arbitary starting values for each of $\theta_2, \dots, \theta_p$ introducing a large amount of error in the early iterations. To avoid this error MCMC generally must be run for a large number of iterations and these early samples are discarded. The computation time per iteration and speed of convergence is extremely problem specific.

We illustrate this with an AR(2), a simple time series model used in the remainder of this section. The AR(2) is described by

\begin{equation}
\label{AR2}
y_t = \mu + \rho_1 y_{t-1} + \rho_2 y_{t-2} + \epsilon_t
\end{equation}

where $\epsilon_t \sim \mathcal{N}(0, \sigma^2)$. We collect the unknown parameters as $\theta = (\mu, \rho_1, \rho_2, \sigma^2)'$ and set priors as

\begin{align}
p(\mu) &\propto 1 \nonumber \\
p(\sigma^2) &\propto \sigma^{-2} \nonumber
\end{align}

add something for $\rho$ to enforce stationarity - maybe switch to an AR(1) or ARMA(1,1).

Describe conditional distributions, possible MH step, result in forecast distribution for $y_{T+1}$

\subsection{Variational Bayes}

Variational Bayes introduces some approximating distribution $q(\theta | \lambda)$ and aims to choose the family $q$ and set of parameters $\lambda$ so that $q(\theta | \lambda$ is as close as possible to the true posterior $p(\theta | y)$. It does this by minimising the Kullback-Leibler (KL) divergence \citep{Kullback1951} from $q(\theta | \lambda)$ to $p(\theta | y)$. The KL divergence is defined by

\begin{equation}
\label{KL-def}
KL[q(\theta | \lambda)||p(\theta | y)] = \int q(\theta | \lambda) \ln \left( \frac{q(\theta | \lambda)}{p(\theta | y)}\right) d\theta.
\end{equation}

and is a non-negative, assymetric measure of the difference between $p(\theta | y)$ and $q(\theta | \lambda)$ that will equal zero if and only if $p(\theta | y) = q(\theta | \lambda)$ almost everywhere \citep{Bishop2006}. It has origins in information theory, and can be interpreted as: \textit{`Given that I know $p(\theta | y)$ for some $\theta \in \Theta$, how much extra information is required, on average, to know the value of $q(\theta | \lambda)$?'} There are examples of the literature of approximations with other measures of divergence, such as \citet{Minka2001} introducting Expectation Propogation (EP), which minimises the reverse measure $KL[p(\theta | y)||q(\theta |\lambda)]$, which was extended to Power-EP in \citet{Minka2004}, which aims to minimise the more general $\alpha-\mbox{divergence}$ \citep{Amari1985}. It is shown in \citet{Bishop2006} (note: find original proof) that minimising $KL[p(\theta | y)||q(\theta | \lambda)]$ used in in EP is equivalent to the MLE of $q(\theta | \lambda)$ given a sample of $\theta$. However, we can write $KL[q(\theta | \lambda)||p(\theta | y)]$ as

\begin{equation}
\label{KL-ELBO}
KL[q(\theta | \lambda)||p(\theta | y)] = \ln(y) - \mathcal{L}(q, y)
\end{equation}

where $\mathcal{L}(q, y)$ is known as the Evidence Lower BOund (ELBO), defined by

\begin{equation}
\label{ELBO}
\mathcal{L}(q, y) = \int_{\theta} q(\theta|\lambda) \ln (p(y, \theta|\lambda)) d\theta -  \int_{\theta} q(\theta|\lambda) \ln (q(\theta|\lambda)) d\theta.
\end{equation}

From (\ref{ELBO}) it is clear that maximising $\mathcal{L}(q, y)$ with respect to $q$ is equivalent to minimising (\ref{KL-def}). Maximising the ELBO is much more convenient than minimising either form of the KL Divergence, and has lead to Variational Bayes been used much more widely in the literature than alternatives such as EP.

\subsubsection{The Mean Field Assumption}

@@

If the likelihood and prior for $\theta_i$ form an exponential family conjugate pair, then this maximisation proceeds by keeping the distributional family constant and simply updating the parameters $\lambda_i$. Otherwise, we can find the optimal distributional family for any $q_i$ by taking the log-joint density and then ignoring any term that doesn't depend on $\theta_i$. Next, take the expectation of any remaining $\theta_{j \neq i}$ and then match to the kernel of a known distribution. If it does not match a known distribution, we may need to make a further approximation $\tilde{q_i(\theta_i|\lambda_i)}$ which has a recognizable distribution. One method as used in \citet{Friston2006} uses a Laplace approximation to and substitute in a Gaussian distribution for an otherwise unrecognizable $q_i(\theta_i | \lambda_i)$.

These steps are almost the same as those used to find a conditional distribution $p(\theta_i | \theta_{j \neq i}, y)$, with dependence on other parameters replaced by their expectations. Hence, the optimal approximating family for $\theta_i$, $q(\theta_i | \lambda_i)$, will come from the same distributional family as the conditional distribution $p(\theta_i | \theta_{j \neq i}, y)$, if it exists in a recognisable form such as the exponential family. Hence deriving the updating algorithm for MFVB takes about the same analytical effort as finding the conditionals for a Gibbs MCMC scheme, and the secondary Laplace approximation will occur in parameters where a Metropolis-Hastings-within-Gibbs step must be used.
\vspace{5mm}

Given these optimal distributional families, each $\lambda_i$ can be found as a function of the data and $\lambda_{j \neq i}$. This dependence requires an algorithm that continuously iterates between each $\lambda_i$ and sets it to its maximising value until $\mathcal{L}(q(\theta | \lambda), y)$ converges with some pre-defined threshold. This is known as a coordinate ascent algorithm and follows below using $k$ as the dimension of $\lambda$.

@@

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Log Joint Density}
 \KwResult{Mean Field Approximation}
 Initialise $\lambda$ randomly\;
 Use (\ref{mf3}) to match each $q(\theta_i|\lambda_i)$ to a tractable distribution\;
 \While{Not converged}{
  \For{$i =  1$ \KwTo $k$}{
      Hold $\lambda_{j \neq i}$ fixed\;
      Update $\lambda_i$ using $y$ and the most recent values of $\lambda_{j \neq i}$\;
     }
 }
 \caption{Coordinate Ascent for MFVB}
  \label{alg:algorithm1}
\end{algorithm}

\subsubsection{Stochastic Variational Bayes}
\emph{Note the change to subsubsection}

In many cases, the posterior is too complex to be captured for a factorising approximation to be reasonabl as we desire our distribution $q(\theta | \lambda)$ to capture dependence between parameters. In this case, the easy maximisation in MFVB is unavailable, and we must resort to what is called Stochastic Variational Bayes (SVB) using a gradient ascent algorithm developed by \citet{Paisley2012} and \citet{Ranganash2014}.

@@

\subsection{Variance Reduction Techniques} \emph{This is general stuff, not just for VB, so can be a subsection}
-Rao Blackwellisation
-Control Variates
-Reparameterisation
\subsection{AR2 model example (Revisited using VB)}

\section{Electricity Load Forecasts}
\subsection{Motivation}
\emph{Will you include a subsection showing an overview of the actual electricity load data you have??}
Real data has five minute updates
Density is of interest due to rare price spikes
Slow convergence of MCMC / Bad scaling to large dimensional models and datasets
\subsection{Exponential Smoothing}
Bayesian Version
Difficulty with smoothing parameters in large models
Infer model structure from draws where possible
\subsection{Variational Bayes Implementation}
Posterior Distributions should not change much as more data is observed, so we can keep approximating distribution family constant
But we might want a model that does change to make updates worthwhile - Markov Switching mechanism to be added?
\subsection{Forecasting}
compare to MCMC when possible

\section{Timeline}
